```{r setup, echo=FALSE, message=FALSE}
## Uncomment after your are ready to publish the post
#opts_knit$set(upload.fun=imgur_upload)
opts_chunk$set(fig.width=5, fig.height=5, cache=TRUE)

## Load knitcitations with a clean bibliography
library(knitcitations)
cleanbib()
cite_options(tooltip=TRUE)

## I made my own citing function since citep() doesn't work like I want to with
## urls that are not really pages themselve like part of a GitHub repo.
mycitep <- function(x, short, year=substr(date(), 21, 24), tooltip=TRUE) {
	tmp <- citep(x)
	res <- gsub("></a>", paste0(">", short, "</a>"), tmp)
	if(tooltip) {
		res <- gsub("\\?\\?\\?\\?", year, res)
	}
	res
}
## Here's an example
# mycitep("https://github.com/lcolladotor/lcollado753", "Collado, 2013")
```
```{r bibsetup, echo=FALSE, message=FALSE, warning=FALSE}
write.bibtex(c("knitcitations" = citation("knitcitations"), "devtools"=citation("devtools")), file = "pkgs.bib")
bib <- read.bibtex("pkgs.bib")
```



Lets say that I want to read in [this R file](https://github.com/lcolladotor/ballgownR-devel/blob/master/ballgownR/R/infoGene.R) from GitHub into R.


The first thing you have to do is locate the raw file. You can do so by clicking on the __Raw__ button in GitHub. In this case it's https://raw.github.com/lcolladotor/ballgownR-devel/master/ballgownR/R/infoGene.R

One would think that using `source()` would work, but it doesn't as shown below:

```{r regular}
source("https://raw.github.com/lcolladotor/ballgownR-devel/master/ballgownR/R/infoGene.R")
```

However, thanks again to Hadley Wickham you can do so by using the `devtools` `r citep(bib[["devtools"]])`
 package.

Here is how it works:
```{r devtools}
library(devtools)
library(roxygen2) 
## Needed because this file has roxygen2 comments. Otherwise you get a 'could not find function "digest"' error
source_url("https://raw.github.com/lcolladotor/ballgownR-devel/master/ballgownR/R/infoGene.R")
## Note that you can specify the SHA-1 hash to be very specific about which version of the file you want to read in.
```

We can then check that `infoGene` has actually been sourced:
```{r check}
"infoGene" %in% ls()
```

That's it!
Enjoy!




Citations made with `knitcitations` `r mycitep(bib[["knitcitations"]], "Boettiger, 2013")`.


```{r bibliography, results='asis', echo=FALSE, cache=FALSE}
## Print bibliography
bibliography()
```

Reproducibility
```{r reproducbility}
sessionInfo()
```
